---
title: "Moving Beyond Linearity"
author: "Carla, Dordor & Hadi"
date: "2/4/2022"
output:
  ioslides_presentation: 
    logo: s.png
    widescreen: yes
    smaller: yes
runtime: shiny
---

```{css, echo=FALSE}
H2 {
  font-size: 24px;
  font-family:calibri;
  color:darkred;
  text-shadow: gray 1px 0 10px;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(prettydoc)
library(caret)
library(MASS)
library(gam)
library(splines)
library(ISLR)
library(Information)
library(sp)
library(data.table)
library(mgcv)
library(gamair)
library(mgcv)
library(data.table)
library(Information)
library(readr)
library(data.table)
data(mcycle)
data(Wage)
data(mpg)

csale<-fread('csale.csv')
new_credit_data<-fread('credit.csv')
dat<-fread('datafortest.csv')
data(meuse)

```

## Fitting linear Model is not good choice!

```{r linear}
renderPlot({
  
  # Fit a linear model
lm_mod <- lm(accel~times, data = mcycle)
# Visualize the model
termplot(lm_mod, partial.resid = TRUE, se = TRUE,col.res = "purple",lwd.term=4,lty.se=2,lwd.se=2)
})
```

```{r include = TRUE}
gam_mod <- gam(accel ~ s(times), data = mcycle, method = "REML")
gam_mod$sp
```

## GAM Flexibility and interpretibility

GAMs....

$$
\sum_{i = 1}^{n}{( y_i - g(x_i))^2} + \lambda\int g^" (t)^2 dt
$$

## Trade off Flexibility VS Overfitting

```{r gam.flexiblity}
inputPanel(
  selectInput("SP", label = "smoothing parameter",
              choices = c(0.1, 0.01, 0.001, 0.0001,0.00001,0.000001), selected = 0.1),
  
  sliderInput("BF", label = "Number of basis functions",
              min = 1, max = 50, value = 1, step = 1)
)

renderPlot({
  gam_mod_s2 <- gam(accel ~ s(times, k = as.numeric(input$BF)), data = mcycle, sp = as.numeric(input$SP))
  
  plot(gam_mod_s2, residuals = TRUE, pch = 1,cex=1,shade = TRUE, shade.col = "hotpink")
})
```

## GAMs for Regression Problem

$$
y_i=\beta_0+\beta_1X_{i1}+\beta_2X_{i2}+...+\beta_pX_{ip}+\epsilon_i
$$ $$
y_i=\beta_0+\sum_{i = 1}^{p}{f_i(X_{ij} )+\epsilon_i} 
$$

$$
= \beta_0+f_1(X_{i1} )+f_2(X_{i2} )+...+f_p(X_{ip} )+\epsilon_i
$$

## GAM for Regression Problem fitting model

$$
FuelConsumption=\beta_0+f_1(weight)+f_2(length)+f_3(price)+\epsilon
$$

```{r gam.regresion}



renderPlot({
      # Fit the model
      mod_city <- gam(city.mpg ~ s(weight)+s(length)+s(price), 
                data = mpg, method = "REML")
      par(mfrow=c(1,3))
      # Plot the model
      plot(mod_city,cex=1,shade = TRUE, shade.col = "gray")
})


```

## GAM for Regression Problem summary

```{r include = TRUE}
mod_city <- gam(city.mpg ~ s(weight)+s(length)+s(price), 
                data = mpg, method = "REML")
summary(mod_city)
```

## GAMs for Classification Problem

$$
log(\frac{p(x)}{1-p(x)})=\beta_0+\beta_1X_1+\beta_2X_2+...+\beta_pX_p
$$

$$
log(\frac{p(x)}{1-p(x)})=\beta_0+f_1(X_1)+f_2(X_2)+...+f_p(X_p)
$$

$$
log(\frac{p(x)}{1-p(x)})=\beta_0+\beta_1\times year+f_2(age)+f_3(education)
$$

## GAMs for Classification Problem

$$
p(x)=Pr(FuelConsumption>24|weight,length,price)
$$

```{r gam.clasification}
renderPlot({
      mod_city_log <- gam(I(city.mpg>24) ~ s(weight)+s(length)+s(price), 
                data = mpg,family=binomial, method = "REML")

      par(mfrow=c(1,3))
      
      plot(mod_city_log,cex=1,shade = TRUE, shade.col = "green")
})


```

## GAMs for Classification Problem summary

```{r include = TRUE}
mod_city_log <- gam(I(city.mpg>24) ~ s(weight)+s(length)+s(price), 
                data = mpg,family=binomial, method = "REML")
summary(mod_city_log)
```

## Adding qualitative (categorical) variable

```{r include = TRUE}
# Fit the model
        mod_city2 <- gam(city.mpg ~ s(weight) +drive, 
                     data = mpg, method = "REML")
summary(mod_city2)
```

## Adding qualitative (categorical) variable plot

```{r outside}
renderPlot({
        # Fit the model
        mod_city2 <- gam(city.mpg ~ s(weight) +drive, 
                     data = mpg, method = "REML")
        # Plot the model
        plot(mod_city2, all.terms = TRUE, pages = 1,cex=1,shade = TRUE, shade.col = "orange")
})
```

## Adding qualitative variable to smooth function {style="font_size:15"}

```{r include = TRUE}
        mod_city3 <- gam(city.mpg ~ s(weight,by=drive)+drive, 
                         data = mpg, method = "REML")
summary(mod_city3)
```

## Adding qualitative variable to smooth function Plot

```{r inside}


renderPlot({
        # Fit the model
        mod_city3 <- gam(city.mpg ~ s(weight,by=drive)+drive,data = mpg, method = "REML")
        par(mfrow=c(1,3))
        # Plot the model
        plot(mod_city3,cex=1,shade = TRUE, shade.col = "lightblue")
})
```

## Seting the number of basis function

```{r }
inputPanel(
  sliderInput("V1", label = "Veriable 1",
              min = 1, max = 10, value = 3, step = 1),
  sliderInput("V2", label = "Veriable 2",
              min = 1, max = 10, value = 3, step = 1),
  sliderInput("V3", label = "Veriable 3",
              min = 1, max = 10, value = 3, step = 1),
  sliderInput("V4", label = "Veriable 4",
              min = 1, max = 10, value = 3, step = 1)
)

renderPlot({
        # Fit the model
mod <- gam(y ~ s(x0, k = as.numeric(input$V1)) + s(x1, k = as.numeric(input$V2)) +
             s(x2, k = as.numeric(input$V3)) + s(x3, k = as.numeric(input$V4)),
           data = dat, method = "REML")

# Check the diagnostics
gam.check(mod)
})
```


## comparing V3_k=3

```{r include = TRUE}
modmod <- gam(y ~ s(x0, k = 3) + s(x1, k = 3) + s(x2, k = 3) + s(x3, k = 3),
           data = dat, method = "REML")
summary(modmod)
```

## comparing V3_k=10

```{r include = TRUE}
modmod2 <- gam(y ~ s(x0, k = 3) + s(x1, k = 3) + s(x2, k = 10) + s(x3, k = 3),
            data = dat, method = "REML")
summary(modmod2)

```

## Checking concurvity on mpg data set

```{r include = TRUE}
mod_con <- gam(hw.mpg ~ s(length) + s(width) + s(height) + s(weight), 
           data = mpg, method = "REML")

concurvity(mod_con,full=FALSE)

```

## 3D Models Using Gam Vis

```{r }
inputPanel(
  sliderInput("C1", label = "layers",
              min = 0, max = 3, value = 0, step = 1),
  sliderInput("C2", label = "rotation",
              min = 0, max = 365, value = 0, step = 1)
)

renderPlot({
# Fit the model
mod2d <- gam(cadmium ~ s(x, y), data = meuse, method = "REML")

vis.gam(mod2d, view = c("x", "y"), 
        plot.type = "persp",se = as.numeric(input$C1),theta = as.numeric(input$C2))
})
```

## HeatMap Using Gam Vis


```{r }
inputPanel(
  sliderInput("B1", label = "Too Far",
              min = 0.05, max = 1, value = 0.05, step = 0.05),
  sliderInput("B2", label = "Levels",
              min = 1, max = 30, value = 1, step = 1)
)

renderPlot({
# Fit the model
mod2d <- gam(cadmium ~ s(x, y), data = meuse, method = "REML")

vis.gam(mod2d, view = c("x", "y"), 
        plot.type = "contour", too.far = as.numeric(input$B1),nlevels=as.numeric(input$B2))
})
```


## fitting clasification model

```{r include=TRUE}
log_mod2 <- gam(purchase ~ s(n_acts) + s(bal_crdt_ratio) + 
                  s(avg_prem_balance) + s(retail_crdt_ratio) +
                  s(avg_fin_balance)  + s(mortgage_age) +
                  s(cred_limit),
                data = csale,
                family = binomial,
                method = "REML")

# View the summary
summary(log_mod2)


```



## prediction and model performance

```{r }
log_mod2 <- gam(purchase ~ s(n_acts) + s(bal_crdt_ratio) + 
                  s(avg_prem_balance) + s(retail_crdt_ratio) +
                  s(avg_fin_balance)  + s(mortgage_age) +
                  s(cred_limit),
                data = csale,
                family = binomial,
                method = "REML")
cred_pp<-predict(log_mod2,newdata=csale, type='response')
csale$predict<- ifelse(cred_pp > 0.5, 1, 0)

confusion_matrix<-table(csale$predict,csale$purchase)

TP<-confusion_matrix[2,2]
FP<-confusion_matrix[2,1]
TN<-confusion_matrix[1,1]
FN<-confusion_matrix[1,2]
#accuracy
accuracy<-(TP + TN) / (TP + FP + TN + FN)
#Errors

err_1<-FP / (TN + FP)
err_2<-FN / (TP + FN)
print(paste('Accuracy is eual to =', accuracy))
print(paste('error Type One is eual to =', err_1))
print(paste('error Type Two is eual to =', err_2))
```



